MS Depression Paper

### 2022-08-13 msdepression/scripts/MS_Depression_First_Paper_started_20220728/msdepression_n380_started_20220728.Rmd
graphics

@@

df_good_mimosa_unique <- df_demo_and_fascicles %>% 
#  filter(depGroupVar != 0) %>%
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup() 
make_demographics_table_ms(data_frame = df_good_mimosa_unique)

@@

# Grouped
barplot_lesion_volume <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
  geom_bar(stat="identity", colour="black", width=0.8, size=1.5) + 
    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
  ggtitle(paste0("Volume of Mimosa Lesions by Diagnosis, p=", round(dep_vs_healthy_t_test_lesion_volume$p.value,3))) +
  ylab("Volume of Lesion") +
  theme(title = element_text(size=14),
        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=20),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        panel.background = element_rect(colour = "white", fill = "white"))# + #comment out if you want pretty 
#  scale_y_continuous(expand = c(0, 0), limits = c(0, 0.52)) #comment out if you want colored graph
print(barplot_lesion_volume) #p=0.067

@@

```{r visualize_depression_network}
##########################################################
#                 Histos                        #
##########################################################

@@

@@

q<-ggplot(lesioned, aes(x=value, fill=depGroupVar)) + geom_histogram() + facet_wrap(~variable)
r<-ggplot(lesioned_dx, aes(x=value, fill=depGroupVar)) + geom_histogram() + facet_wrap(~variable)
r_color<-ggplot(lesioned_dx, aes(x=value, color=factor(depGroupVar))) + geom_histogram() + facet_wrap(~variable)
x<-ggplot(lesioned, aes(x=value)) + geom_histogram() + facet_wrap(~variable)
y<-ggplot(healthy, aes(x=value)) + geom_histogram() + facet_wrap(~variable)
z<-ggplot(depressed, aes(x=value)) + geom_histogram() + facet_wrap(~variable)

@@

#### look at mean fascicle volume across participants and transpose
# drop exam date and empi, and transpose
fascicles_alone_df <- t(added_dep_to_fascicles[,!(colnames(added_dep_to_fascicles) %in% c("EMPI", "EXAM_DATE", "depGroupVar"))])
fascicle_names <- rownames(fascicles_alone_df)

@@

@@

sorted_ggplot<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value))+
  ggtitle("Proportion of Volume per Fascicle Affected") + 
  ylab("Proportion") + 
  geom_bar(stat="identity") +
  ylim(0,0.8) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.title = element_text(hjust = 0.5, size=14),
        axis.title.x = element_text(size=14),
        axis.title.y = element_text(size=14))
print(sorted_ggplot)

######### 
#############
#### look at mean fascicle volume across healthy participants and transpose
#############

@@

@@

# ordered by mean value overall
fascicle_values_sorted_by_overlap$fascicle <- factor(fascicle_values_sorted_by_overlap$fascicle, levels = order_of_mean_values_overall)
sorted_ggplot_healthy_ordered<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value))+
  ggtitle("Proportion of Volume per Fascicle Affected Ordered by mean values NonDepressed") + 
  geom_bar(stat="identity") +
  ylab("Proportion") +
  ylim(0,0.8) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.title = element_text(hjust = 0.5, size=14),
        axis.title.x = element_text(size=14),
        axis.title.y = element_text(size=14))
print(sorted_ggplot_healthy_ordered)

@@

@@

# ordered by mean value overall
fascicle_values_sorted_by_overlap$fascicle <- factor(fascicle_values_sorted_by_overlap$fascicle, levels = order_of_mean_values_overall)
sorted_ggplot_depressed_ordered<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value))+
  ggtitle("Proportion of Volume per Fascicle Affected Ordered by mean values Depressed") + 
  ylab("Proportion Affected") +
  geom_bar(stat="identity") +
  ylim(0,0.8) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.title = element_text(hjust = 0.5, size=14),
        axis.title.x = element_text(size=14),
        axis.title.y = element_text(size=14))
print(sorted_ggplot_depressed_ordered)

@@

plot_all_fascicles_r_and_l_sep_top_quartile <-ggplot(data = fascicles_dep_25_percent_for_plot, aes(x = fascicle, y=value))+
  ggtitle("Proportion of Volume per Fascicle Affected Dep net, top quartile") + 
  geom_bar(stat="identity") +
  ylab("Proportion") +
  ylim(0,0.8) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.title = element_text(hjust = 0.5, size=14),
        axis.title.x = element_text(size=14),
        axis.title.y = element_text(size=14))
print(plot_all_fascicles_r_and_l_sep_top_quartile)

@@

plot_all_fascicles_r_and_l_sep_bottom_3_quartiles <-ggplot(data = fascicles_dep_bottom_3_quartiles_for_plot, aes(x = fascicle, y=value))+
  ggtitle("Proportion of Volume per Fascicle Affected, bottom three quartiles") + 
  ylab("Proportion") +
  geom_bar(stat="identity") +
  ylim(0,0.8) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.title = element_text(hjust = 0.5, size=14),
        axis.title.x = element_text(size=14),
        axis.title.y = element_text(size=14))
print(plot_all_fascicles_r_and_l_sep_bottom_3_quartiles)

>>>>

#### Using volume rather than proportion to determine inside vs outside depression network
It occurred to me that defining the top 25% depression network has been done by looking at proportion overlap, where everything else seems to work best in volume. It made me wonder what would happen if I defined my top 25% network based on the absolute volume of overlap a fascicle had with the depression network, and ranked based on that (rather than the % of fascicle that overlapped). It tests a slightly different hypothesis, where the depression network is though of as a collection of fascicles that have the most dead ending or overlap, i.e. contribute the most volume to the network. Proportion, on the other hand, thinks more about the specificity of the fascicle - i.e. if the majority of a fascicle dead ends in the network, its activity is probably highly specific to the network. In a volume only, the value of that fascicle's contribution would be neglibile. Using volume = house of representatives. using overlap - senate.

Or put another way, if you do by volume, big fascicles get the most potential votes, by virtue of their bigness. It would be entirely possible for the AF, for example, to have some of the highest overlap with respect to volume, but if that volume is only 10% of the AF fascicle, it means that 90% dead end outside of the depression network. If a tiny fascicle is entirely hit, at least we know there is some specificity to its existence. Conversely, if the AF accounts for 99% of the volume overlap w/the depression network, even if it only 10% of the volume of the AF, it seems like it should be valued highly as part of the depression network. Luckily, results from both work out well.

Results after defining the depresion 25% network based on the volume rather than proportion overlap: i.e. a fascicle is in the "depression network" if the volume that it shares with the depression network is in the top 25%. 
  - Irrespective of network, overall disease by volume: Dep > nondep (p=0.045). 
  - Within dep network: dep disease > nondep disease i, p=0.0.034,
  - Outside dep network (bottom 75%), dep disease ~= nondep disease, p = 0.071 (NS)

dep_network_names_top_quartile_no_cranial_nerves_by_vol
 [1] CC      AF_L    SLF2_R  SLF3_R  FAT_L   TR_S_R  SLF2_L  AF_R    TR_S_L  FAT_R   CS_S_R  CPT_F_R CPT_P_L ILF_R   CPT_P_R CS_S_L  PAT_R   MdLF_L  CBT_R

dep_network_names_bottom_3_quartiles_no_cranial_nerves_by_vol
  [1] C_FPH_L C_FPH_R C_FP_L  C_FP_R  C_PH_L  C_PHP_L C_PHP_R C_PH_R  C_R_L   C_R_R   EMC_L   EMC_R   IFOF_L  IFOF_R  ILF_L   MdLF_R  PAT_L   SLF1_L  SLF1_R 
[20] SLF3_L  UF_L    UF_R    VOF_L   VOF_R   CB_L    CB_R    ICP_L   ICP_R   MCP     SCP     V       AR_L    AR_R    CBT_L   CPT_F_L CPT_O_L CPT_O_R CS_A_L 
[39] CS_A_R  CS_P_L  CS_P_R  CST_L   CST_R   DRTT_L  DRTT_R  F_L     F_R     ML_L    ML_R    OR_L    OR_R    RST_L   RST_R   TR_A_L  TR_A_R  TR_P_L  TR_P_R 
[58] AC

@@

df_good_mimosa_unique <- df_demo_and_fascicles %>% 
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup() 
make_demographics_table_ms(data_frame = df_good_mimosa_unique)

df_good_mimosa_unique_dep_and_healthy <- df_demo_and_fascicles %>% 
  filter(depGroupVar != 0) %>%
  group_by(EMPI) %>% 
  arrange(EXAM_DATE) %>% 
  slice(1) %>% 
  ungroup() 
make_demographics_table_ms(data_frame = df_good_mimosa_unique_dep_and_healthy)

@@

# Grouped
barplot_lesion_volume <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
  geom_bar(stat="identity", colour="black", width=0.7, size=1.5) + 
    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
  ggtitle(paste0("Volume of Mimosa Lesions by Diagnosis, p=", round(dep_vs_healthy_t_test_lesion_volume$p.value,3))) +
  ylab("Volume of Lesion") +
  theme(title = element_text(size=16),
        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=20),
        aspect.ratio = 5/5,
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        panel.background = element_rect(colour = "white", fill = "white"))# + #comment out if you want pretty 
#  scale_y_continuous(expand = c(0, 0), limits = c(0, 0.52)) #comment out if you want colored graph
print(barplot_lesion_volume) #p=0.067

barplot_lesion_volume_range <- ggplot(data, aes(x=reorder(Diagnosis, -value),fill=Diagnosis, y=value)) + 
  geom_bar(stat="identity", colour="black", width=0.7, size=1.5) + 
    geom_errorbar(aes(ymin=value-sem, ymax=value+sem), width=.1, position=position_dodge(0.9), size=1.5) +
  scale_fill_manual(values = c("#cdcdcd", "#FFFFFF")) + #comment out if you want colored graph
  ggtitle(paste0("Volume of Mimosa Lesions by Diagnosis, p=", round(dep_vs_healthy_t_test_lesion_volume$p.value,3))) +
  ylab("Volume of Lesion") +
  ylim(0,20000) +
  theme(title = element_text(size=16, hjust = 0.2),
        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=20),
        aspect.ratio = 5/5,
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        panel.background = element_rect(colour = "white", fill = "white"))# + #comment out if you want pretty 
 # scale_y_continuous(expand = c(0, 0), limits = c(0, 20000)) #comment out if you want colored graph
print(barplot_lesion_volume_range) #p=0.067

##### violin plot

#extract overall volume and diagnosis
df_lesion_volume_and_diagnosis <- dep_and_healthy_groups_for_ICD_analysis %>% 
  dplyr::select(volume_of_mimosa_lesions,depGroupVar) %>%
  filter(depGroupVar != 0) %>% 
  mutate(depGroupVar = ifelse(depGroupVar == 1, "Non-depressed", "Depressed"))

violin_lesion_volume <- ggplot(df_lesion_volume_and_diagnosis, aes(x=depGroupVar, y=volume_of_mimosa_lesions, color=depGroupVar)) +#, fill=depGroupVar)) + 
  geom_violin() + #draw_quantiles = c(0.25, 0.5, 0.75)) + 
  stat_summary(fun= "mean", geom = "pointrange", colour = "black") +
  geom_boxplot(width=0.1) + theme_minimal() + 
  scale_x_discrete(labels=c("1" = "Depressed", "2" = "Non-Depressed")) + 
  ggtitle(paste0("Volume of Lesions by Diagnosis")) +
  ylab("Volume of Lesions") +
  theme(plot.title = element_text(size=16, hjust = 0.8),
        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=16),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        legend.position = "none",
        aspect.ratio=7/5,
        panel.background = element_rect(colour = "white", fill = "white"))# + #comment out if you want pretty 
print(violin_lesion_volume) #p=0.067

jitter_lesion_volume <- ggplot(df_lesion_volume_and_diagnosis, aes(x=depGroupVar, y=volume_of_mimosa_lesions, color=depGroupVar)) +#, fill=depGroupVar)) + 
  geom_jitter(shape=16, position=position_jitter(0.15)) + 
  geom_boxplot(width=0.1, color="black", fill="white", outlier.size = 0.2) + #theme_minimal(color="black") + 
  scale_x_discrete(labels=c("1" = "Depressed", "2" = "Non-Depressed")) + 
  ggtitle(paste0("Volume of Lesions by Diagnosis")) +
  ylab("Volume of Lesions") +
  theme(plot.title = element_text(size=16, hjust = 0.8),
        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=16),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        legend.position = "none",
        aspect.ratio=7/5,
        panel.background = element_rect(colour = "white", fill = "white"))# + #comment out if you want pretty 
print(jitter_lesion_volume) #p=0.067

violin_lesion_volume <- ggplot(df_lesion_volume_and_diagnosis, aes(x=depGroupVar, y=volume_of_mimosa_lesions, fill=depGroupVar)) +#, fill=depGroupVar)) + 
  geom_violin() + 
  geom_boxplot(width=0.1, color="black", fill="white", outlier.size = 0.2) + #theme_minimal(color="black") + 
  scale_x_discrete(labels=c("1" = "Depressed", "2" = "Non-Depressed")) + 
  ggtitle(paste0("Volume of Lesions by Diagnosis")) +
  ylab("Volume of Lesions") +
  theme(plot.title = element_text(size=16, hjust = 0.8),
        axis.text = element_text(size = 14, colour = "black"), axis.title.x = element_blank(), 
        axis.title.y = element_text(size=16),
        axis.line = element_line(size = 1.5, colour = "black", linetype=1),
        legend.position = "none",
        aspect.ratio=7/5,
        panel.background = element_rect(colour = "white", fill = "white"))# + #comment out if you want pretty 
print(violin_lesion_volume) #p=0.067

@@

```{r visualize_depression_network_proportion}
##########################################################
#                 Histos                        #
##########################################################

@@

```{r visualize_depression_network_volume}
##########################################################
#                 Histos                        #
##########################################################

######## Histograms for depression network
#sort based on overlap
fascicle_values_dep_sorted_by_overlap <- fascicle_volumes_dep
fascicle_values_dep_sorted_by_overlap$non_zero_voxels_in_dep_map[is.na(fascicle_values_dep_sorted_by_overlap$non_zero_voxels_in_dep_map)] = 0
fascicle_values_dep_sorted_by_overlap$fascicle <- factor(fascicle_values_dep_sorted_by_overlap$fascicle, levels = fascicle_values_dep_sorted_by_overlap$fascicle[order(fascicle_values_dep_sorted_by_overlap$non_zero_voxels_in_dep_map)])

plot_all_fascicles_r_and_l_sep_no_cranial_nerves <-ggplot(data = fascicles_r_and_l_sep_no_cranial_nerves, aes(x = fascicle, y=non_zero_voxels_in_dep_map))+
  ggtitle("Volume Overlap per Fascicle 3_09") + 
  geom_bar(stat="identity", fill="006300") +
  xlab("Fascicle") +
  ylab("Overlap with Depression Mask") +
  ylim(0,40000) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.title = element_text(hjust = 0.5, size=14),
        axis.title.x = element_text(size=14),
        axis.title.y = element_text(size=14),
        axis.line = element_line(size=1),
        panel.background = element_rect(colour = "white", fill = "white"))
print(plot_all_fascicles_r_and_l_sep_no_cranial_nerves)

#sort all the fascicles by %overlap, in decreasing order (so highest overlap at the top), and then take the top 1-quartile
dep_network_names_top_quartile_no_cranial_nerves_by_vol<- (fascicle_values_dep_no_cranial_nerves[order(fascicle_values_dep_no_cranial_nerves$non_zero_voxels_in_dep_map, decreasing = TRUE),])$fascicle[1:quartile] #n=19

#get a network that includes all of the fascicles NOT in dep 25% network
dep_network_names_bottom_3_quartiles_no_cranial_nerves_by_vol <- fascicle_values_dep_no_cranial_nerves$fascicle[!(fascicle_values_dep_no_cranial_nerves$fascicle %in% dep_network_names_top_quartile_no_cranial_nerves_by_vol)] #n=68

fascicles_dep_25_percent_for_plot_no_cranial_nerves <- fascicle_values_dep_sorted_by_overlap[(fascicle_values_dep_sorted_by_overlap$fascicle %in% dep_network_names_top_quartile_no_cranial_nerves_by_vol),]

plot_all_fascicles_r_and_l_sep_top_quartile_no_cranial_nerves <-ggplot(data = fascicles_dep_25_percent_for_plot_no_cranial_nerves, aes(x = fascicle, y=non_zero_voxels_in_dep_map)) + 
  geom_bar(stat="identity", fill="violet") +
  ggtitle("Volume per Fascicle Affected Dep net 3_09, top quartile") + 
  ylab("Volume in Depression Mask") +
  ylim(0,40000) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.title = element_text(hjust = 0.5, size=14),
        axis.title.x = element_text(size=14),
        axis.title.y = element_text(size=14),
        axis.line = element_line(size=1),
        panel.background = element_rect(colour = "white", fill = "white"))
print(plot_all_fascicles_r_and_l_sep_top_quartile_no_cranial_nerves)

#bottom 3 quartiles
fascicles_dep_bottom_3_quartiles_for_plot_no_cranial_nerves<- fascicle_values_dep_sorted_by_overlap[(fascicle_values_dep_sorted_by_overlap$fascicle %in% dep_network_names_bottom_3_quartiles_no_cranial_nerves_by_vol),]

plot_all_fascicles_r_and_l_sep_bottom_3_quartiles_no_cranial_nerves <-ggplot(data = fascicles_dep_bottom_3_quartiles_for_plot_no_cranial_nerves, aes(x = fascicle, y=non_zero_voxels_in_dep_map)) + 
  geom_bar(stat="identity", fill="darkgreen") +
  ylab("Volume in Depression Mask") +
  ylim(0,40000) +
  ggtitle("Volume per Fascicle Affected Dep net 3_09, bottom three quartiles") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.title = element_text(hjust = 0.5),
        axis.line = element_line(size=1),
        panel.background = element_rect(colour = "white", fill = "white"))
print(plot_all_fascicles_r_and_l_sep_bottom_3_quartiles_no_cranial_nerves) #n=58, which is total 68 - 10 CNs

@@

q<-ggplot(lesioned, aes(x=value, fill=depGroupVar)) + geom_histogram() + facet_wrap(~variable)
r<-ggplot(lesioned_dx, aes(x=value, fill=depGroupVar)) + geom_histogram() + facet_wrap(~variable)
r_color<-ggplot(lesioned_dx, aes(x=value, color=factor(depGroupVar))) + 
  geom_histogram(color="darkgreen") + 
  facet_wrap(~variable)
x<-ggplot(lesioned, aes(x=value)) + geom_histogram() + facet_wrap(~variable)
y<-ggplot(healthy, aes(x=value)) + geom_histogram() + facet_wrap(~variable)
z<-ggplot(depressed, aes(x=value)) + geom_histogram() + facet_wrap(~variable)

@@

#### look at mean fascicle volume overlap across participants and transpose
# drop exam date and empi, and transpose
fascicles_alone_df <- t(added_dep_to_fascicles[,!(colnames(added_dep_to_fascicles) %in% c("EMPI", "EXAM_DATE", "depGroupVar"))])
fascicle_names <- rownames(fascicles_alone_df)

@@

write.table(fascicles_alone_rowMeans, file = paste0(homedir, "/results/fascicle_mean_proportion_overlap_nfasc_77.csv"), row.names = FALSE, col.names = FALSE, quote=FALSE, sep = ",")

@@

sorted_ggplot<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value, fill=value))+
  ggtitle("Proportion of Volume per Fascicle Affected") + 
  ylab("Proportion") + 
  geom_bar(stat="identity") + 
 # scale_fill_continuous(type="viridis") + 
  scale_fill_gradient(low = "red", high = "yellow") +
  ylim(0,0.8) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.title = element_text(hjust = 0.5, size=14),
        axis.title.x = element_text(size=14),
        axis.title.y = element_text(size=14),
        axis.line = element_line(size=1),
        panel.background = element_rect(colour = "white", fill = "white"))
print(sorted_ggplot)

######### 
#############
#### look at mean fascicle overlap across healthy participants and transpose
#############

@@

write.table(fascicles_alone_rowMeans, file = paste0(homedir, "/results/fascicle_mean_proportion_overlap_nfasc_77_nondep.csv"), row.names = FALSE, col.names = FALSE, quote=FALSE, sep = ",")

@@

# ordered by mean value overall
fascicle_values_sorted_by_overlap$fascicle <- factor(fascicle_values_sorted_by_overlap$fascicle, levels = order_of_mean_values_overall)
sorted_ggplot_healthy_ordered<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value, fill=value))+
  ggtitle("Proportion of Volume per Fascicle Affected Ordered by mean values NonDepressed") + 
  geom_bar(stat="identity") + 
  scale_fill_continuous(type="viridis") +
  #scale_fill_gradient(low = "red", high = "yellow") +
  ylab("Proportion Affected") +
  ylim(0,0.8) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.title = element_text(hjust = 0.5, size=14),
        axis.title.x = element_text(size=14),
        axis.title.y = element_text(size=14),
        axis.line = element_line(size=1),
        panel.background = element_rect(colour = "white", fill = "white"))
print(sorted_ggplot_healthy_ordered)

@@

write.table(fascicles_alone_rowMeans, file = paste0(homedir, "/results/fascicle_mean_proportion_overlap_nfasc_77_dep.csv"), row.names = FALSE, col.names = FALSE, quote=FALSE, sep = ",")

@@

# ordered by mean value overall
fascicle_values_sorted_by_overlap$fascicle <- factor(fascicle_values_sorted_by_overlap$fascicle, levels = order_of_mean_values_overall)
sorted_ggplot_depressed_ordered<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value, fill=value))+
  ggtitle("Proportion of Volume per Fascicle Affected Ordered by mean values Depressed") + 
  ylab("Proportion Affected") +
  geom_bar(stat="identity") + 
  #scale_fill_gradient(low = "red", high = "yellow") +
  scale_fill_continuous(type="viridis") + #, color="orange") +
  ylim(0,0.8) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.title = element_text(hjust = 0.5, size=14),
        axis.title.x = element_text(size=14),
        axis.title.y = element_text(size=14),
        axis.line = element_line(size=1),
        panel.background = element_rect(colour = "white", fill = "white"))
print(sorted_ggplot_depressed_ordered)

### color by quartile
#num_buckets = round(length(fascicle_values_sorted_by_overlap$fascicle)/4,0) #get the number of quartiles
#color_columns= rep(x = c("#8DD3C7", "#FFFFB3", "#BEBADA", "#FB8072"), times=1, each=num_buckets) #assign each quartile a color
#color_columns= rep(x = c("violet", "blue", "green", "red"), times=1, each=num_buckets) #assign each quartile a color

#if (length(fascicle_values_sorted_by_overlap$fascicle) %% 4 != 0){
 # color_columns=c("#8DD3C7", color_columns) #add one more to the beginning of the vector since the number is odd, so mod is odd number
#}

fascicle_values_sorted_by_overlap$fascicle <- factor(fascicle_values_sorted_by_overlap$fascicle, levels = order_of_mean_values_overall)
sorted_ggplot_depressed_ordered<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value, fill=value))+
  ggtitle("Proportion of Volume per Fascicle Affected Ordered by mean values Depressed") + 
  geom_bar(stat="identity") + 
  scale_fill_continuous(type = "viridis") + 
  ylab("Proportion Affected") +
  ylim(0,0.8) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.title = element_text(hjust = 0.5, size=14),
        axis.title.x = element_text(size=14),
        axis.title.y = element_text(size=14),
        axis.line = element_line(size=1),
        panel.background = element_rect(colour = "white", fill = "white"))
print(sorted_ggplot_depressed_ordered)

@@

plot_all_fascicles_r_and_l_sep_top_quartile <-ggplot(data = fascicles_dep_25_percent_for_plot, aes(x = fascicle, y=value))+
  ggtitle("Proportion of Volume per Fascicle Affected Dep net, top quartile") + 
  geom_bar(stat="identity", fill="darkblue") +
  ylab("Proportion") +
  ylim(0,0.8) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.title = element_text(hjust = 0.5, size=14),
        axis.title.x = element_text(size=14),
        axis.title.y = element_text(size=14),
        axis.line = element_line(size=1),
        panel.background = element_rect(colour = "white", fill = "white"))
print(plot_all_fascicles_r_and_l_sep_top_quartile)

@@

plot_all_fascicles_r_and_l_sep_bottom_3_quartiles <-ggplot(data = fascicles_dep_bottom_3_quartiles_for_plot, aes(x = fascicle, y=value))+
  ggtitle("Proportion of Volume per Fascicle Affected, bottom three quartiles") + 
  ylab("Proportion") +
  geom_bar(stat="identity", fill="red") +
  ylim(0,0.8) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.title = element_text(hjust = 0.5, size=14),
        axis.title.x = element_text(size=14),
        axis.title.y = element_text(size=14),
        axis.line = element_line(size=1),
        panel.background = element_rect(colour = "white", fill = "white"))
print(plot_all_fascicles_r_and_l_sep_bottom_3_quartiles)

############ Look at mean overall volume across all brain regions
#### look at mean fascicle volume overlap across participants and transpose
# drop exam date and empi, and transpose
fascicles_alone_df_vol <- t(df_demo_and_fascicle_volumes_full[,(colnames(df_demo_and_fascicle_volumes_full) %in% fascicle_names_no_cranial_nerves)])
fascicle_names <- rownames(fascicles_alone_df_vol)

#take row means
fascicles_alone_vol_rowMeans <- data.frame(fascicle_names, rowMeans(fascicles_alone_df_vol))
names(fascicles_alone_vol_rowMeans) <- c("fascicle", "value")

#GET RID OF CRANIAL NERVES
fascicles_alone_vol_rowMeans <- fascicles_alone_vol_rowMeans %>% filter(fascicle %in% fascicle_names_no_cranial_nerves)

write.table(fascicles_alone_vol_rowMeans, file = paste0(homedir, "/results/fascicle_mean_volume_nfasc_77.csv"), row.names = FALSE, col.names = FALSE, quote=FALSE, sep = ",")

#sort based on overlap
fascicle_values_sorted_by_overlap <- fascicles_alone_vol_rowMeans

######## Histograms/bargraphs
#this histogram helps contextualize the number of streamlines with varying degrees of proportions affected
hist(x=fascicle_values_dep_sorted_by_overlap$prop_in_mask, main = "Histogram Affected Streamlines", xlab = "Volume per Fascicle Affected", breaks = 20)

sorted_ggplot<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value, fill=value))+
  ggtitle("Volume per Fascicle Affected") + 
  ylab("Volume") + 
  geom_bar(stat="identity") + 
  scale_fill_continuous(type="viridis") + 
  ylim(0,230000) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.title = element_text(hjust = 0.5, size=14),
        axis.title.x = element_text(size=14),
        axis.title.y = element_text(size=14),
        axis.line = element_line(size=1),
        panel.background = element_rect(colour = "white", fill = "white"))
print(sorted_ggplot)

sorted_ggplot<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value, fill=value))+
  ggtitle("Volume per Fascicle Affected, exclude CC") + 
  ylab("Volume") + 
  geom_bar(stat="identity") + 
  scale_fill_continuous(type="viridis") + 
  ylim(0,30000) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.title = element_text(hjust = 0.5, size=14),
        axis.title.x = element_text(size=14),
        axis.title.y = element_text(size=14),
        axis.line = element_line(size=1),
        panel.background = element_rect(colour = "white", fill = "white"))
print(sorted_ggplot)

##### healthy volume
df_demo_and_fascicle_volumes_full_nondep <- df_demo_and_fascicle_volumes_full %>% filter(depGroupVar==1)

# drop exam date and empi, and transpose
fascicles_alone_df_vol <- t(df_demo_and_fascicle_volumes_full_nondep[,(colnames(df_demo_and_fascicle_volumes_full_nondep) %in% fascicle_names_no_cranial_nerves)])
fascicle_names <- rownames(fascicles_alone_df_vol)

#take row means
fascicles_alone_vol_rowMeans <- data.frame(fascicle_names, rowMeans(fascicles_alone_df_vol))
names(fascicles_alone_vol_rowMeans) <- c("fascicle", "value")

#GET RID OF CRANIAL NERVES
fascicles_alone_vol_rowMeans <- fascicles_alone_vol_rowMeans %>% filter(fascicle %in% fascicle_names_no_cranial_nerves)

write.table(fascicles_alone_vol_rowMeans, file = paste0(homedir, "/results/fascicle_mean_volume_nfasc_77_nondep.csv"), row.names = FALSE, col.names = FALSE, quote=FALSE, sep = ",")

#sort based on overlap
fascicle_values_sorted_by_overlap <- fascicles_alone_vol_rowMeans

######## Histograms/bargraphs
#this histogram helps contextualize the number of streamlines with varying degrees of proportions affected
hist(x=fascicle_values_dep_sorted_by_overlap$prop_in_mask, main = "Histogram Affected Streamlines", xlab = "Volume per Fascicle Affected", breaks = 20)

sorted_ggplot<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value, fill=value))+
  ggtitle("Volume per Fascicle Affected Nondepressed") + 
  ylab("Volume") + 
  geom_bar(stat="identity") + 
  scale_fill_continuous(type="viridis") + 
  ylim(0,230000) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.title = element_text(hjust = 0.5, size=14),
        axis.title.x = element_text(size=14),
        axis.title.y = element_text(size=14),
        axis.line = element_line(size=1),
        panel.background = element_rect(colour = "white", fill = "white"))
print(sorted_ggplot)

sorted_ggplot<-ggplot(data = fascicle_values_sorted_by_overlap, aes(x = fascicle, y=value, fill=value))+
  ggtitle("Volume per Fascicle Affected Nondepressed, exclude CC") + 
  ylab("Volume") + 
  geom_bar(stat="identity") + 
  scale_fill_continuous(type="viridis") + 
  ylim(0,30000) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        plot.title = element_text(hjust = 0.5, size=14),
        axis.title.x = element_text(size=14),
        axis.title.y = element_text(size=14),
        axis.line = element_line(size=1),
        panel.background = element_rect(colour = "white", fill = "white"))
print(sorted_ggplot)

##### depressed volume
# drop exam date and empi, and transpose

df_demo_and_fascicle_volumes_full_dep <- df_demo_and_fascicle_volumes_full %>% filter(depGroupVar==2)

fascicles_alone_df_vol <- t(df_demo_and_fascicle_volumes_full_dep[,(colnames(df_demo_and_fascicle_volumes_full_dep) %in% fascicle_names_no_cranial_nerves)])
fascicle_names <- rownames(fascicles_alone_df_vol)

#take row means
fascicles_alone_vol_rowMeans <- data.frame(fascicle_names, rowMeans(fascicles_alone_df_vol))
names(fascicles_alone_vol_rowMeans) <- c("fascicle", "value")

#GET RID OF CRANIAL NERVES
fascicles_alone_vol_rowMeans <- fascicles_alone_vol_rowMeans %>% filter(fascicle %in% fascicle_names_no_cranial_nerves)

write.table(fascicles_alone_vol_rowMeans, file = paste0(homedir, "/results/fascicle_mean_volume_nfasc_77_dep.csv"), row.names = FALSE, col.names = FALSE, quote=FALSE, sep = ",")

#sort based on overlap
fascicle_values_sorted_by_overlap <- fascicles_alone_vol_rowMeans

######## Histograms/bargraphs
#this histogram helps contextualize the number of streamlines with varying degrees of proportions affected
hist(x=fascicle_values_dep_sorted_by_overlap$prop_in_mask, main = "Histogram Affected Streamlines", xlab = "Volume per Fascicle Affected", breaks = 20)
